---
title: "DESeq on Pauline data"
author: "Pauline"
date: "6 juin 2018"
output: html_document
---
Zebrafish caudal fin regeneration in adults vs fin fold regeneration in zebrafish larvae

# Info about raw data

```{r, echo=F, message=F, warning=F}
rm(list=ls())
## Path to the input files
indir="J:\\Pauline J\\Pauline"
## Chez Marie
indir="/home/msemon/Documents/ProjetPauline/Fin_regeneration_project"

## Path to the output directory
outdir="J:\\Pauline J\\Pauline\\DESeq"
## Chez Marie
outdir="/home/msemon/Documents/ProjetPauline/Fin_regeneration_project"

## Name of read count file
count_file="counts_raw.txt"

## Name of the metadata file
MD="metadata.txt"

## Name of the annotation file
annot.file="annotation.txt"



### File importation

setwd(indir)
library(data.table, quietly = T, warn.conflicts = F)
rawcounts=fread(count_file, data.table = F)
metadata=read.table(MD, header=T)

### Expressed genes

print(paste("Read counts for", nrow(rawcounts), "genes", sep=" "))

rownames(rawcounts)=rawcounts$Gene_ID
rawcounts=rawcounts[,-1]

w=which(rowSums(rawcounts)==0)
print(paste(length(w), "genes are expressed in none of the samples"))

```

# DESeq on adult samples

```{r, message=F, warning=F}
metaAdult=metadata[metadata$Stade=="Adult",]
count=rawcounts[,match(metaAdult$Bam_File_Name, colnames(rawcounts))]

info=metaAdult[,4:5]
rownames(info)=colnames(count) 

library(DESeq2, quietly = T, warn.conflicts = F)
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)

dds0 <- estimateSizeFactors(dds)

log.norm.counts <- log2(counts(dds0, normalized=TRUE) + 1)
norm.counts <- counts(dds0, normalized=TRUE)
#setwd(outdir)
#write.table(norm.counts, "Adult_DESeq_normalized_data.txt", sep="\t", quote=F, dec=",")

normPCA=t(norm.counts)
normPCA=data.frame(metaAdult[,4:5], normPCA, row.names = metaAdult$Description)
library(FactoMineR, quietly = T, warn.conflicts = F)
res.pca=PCA(normPCA, quali.sup = 1:2, graph=F)
plot.PCA(res.pca, choix="ind", habillage=1, axes=c(1,2), label = "quali")
plot.PCA(res.pca, choix="ind", habillage=2, axes=c(1,2), label = "quali")

dd=dimdesc(res.pca)

#Correlation of illustrative variables to the PCA axe 1
dd$Dim.1$quali
#Correlation of illustrative variables to the PCA axe 2
dd$Dim.2$quali
```

--> The 1st and 2nd axis of the PCA represent about 45% of the total variation and are correlated with the condition.

--> The first axis separate the ablated samples from the samples before ablation and complete regeneration. 

## 2dpa vs 0dpa

```{r, echo=T, message=F, warning=F}
dds0 <- DESeq(dds0)
#resultsNames(dds0)
res20 <- results(dds0, name="Condition_2dpa_vs_0dpa")
#Number of DE genes :
table(res20$padj < 0.05)
```

## 3dpa vs 2dpa

```{r, echo=T, message=F, warning=F}
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)
dds$Condition <- relevel(dds$Condition, ref = "2dpa")

dds2 <- estimateSizeFactors(dds)
dds2 <- DESeq(dds2)
#resultsNames(dds2)
res23 <- results(dds2, name="Condition_3dpa_vs_2dpa")
print("Number of DE genes :")
table(res23$padj < 0.05)
```


## 10dpa vs 3dpa

```{r, echo=F, message=F, warning=F}
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)
dds$Condition <- relevel(dds$Condition, ref = "3dpa")

dds3 <- estimateSizeFactors(dds)
dds3 <- DESeq(dds3)
#resultsNames(dds3)
res23 <- results(dds3, name="Condition_10dpa_vs_3dpa")
print("Number of DE genes :")
table(res23$padj < 0.05)
```

## 0dpa vs 10dpa

```{r, echo=F, message=F, warning=F}
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)
dds$Condition <- relevel(dds$Condition, ref = "10dpa")

dds10 <- estimateSizeFactors(dds)
dds10 <- DESeq(dds10)
#resultsNames(dds10)
res010 <- results(dds10, name="Condition_0dpa_vs_10dpa")
print("Number of DE genes :")
table(res010$padj < 0.05)
```