---
title: "DESeq on Pauline data"
author: "Pauline"
date: "6 juin 2018"
output: html_document
---
Zebrafish caudal fin regeneration in adults vs fin fold regeneration in zebrafish larvae

# Info about raw data

```{r, echo=F, message=F, warning=F}
rm(list=ls())
## Path to the input files
indir="J:\\Pauline J\\Pauline\\Fin_regeneration_project/"
## Chez Marie
#indir="/home/msemon/Documents/ProjetPauline/Fin_regeneration_project"

## Path to the output directory
outdir="J:\\Pauline J\\Pauline\\Fin_regeneration_project\\DESeq"
## Chez Marie
#outdir="/home/msemon/Documents/ProjetPauline/Fin_regeneration_project"

## Name of read count file
count_file="counts_raw.txt"

## Name of the metadata file
MD="metadata.txt"

## Name of the annotation file
annot.file="annotation.txt"



### File importation

setwd(indir)
library(data.table, quietly = T, warn.conflicts = F)
rawcounts=fread(count_file, data.table = F)
metadata=read.table(MD, header=T)
metadata$Description=gsub("-", "_", metadata$Description)

### Expressed genes

print(paste("Read counts for", nrow(rawcounts), "genes", sep=" "))

rownames(rawcounts)=rawcounts$Gene_ID
rawcounts=rawcounts[,-1]

w=which(rowSums(rawcounts)==0)
print(paste(length(w), "genes are expressed in none of the samples"))

```

# DESeq on adult samples

```{r, message=F, warning=F}
metaAdult=metadata[metadata$Stade=="Adult",]
count=rawcounts[,match(metaAdult$Bam_File_Name, colnames(rawcounts))]

info=metaAdult[,c(4,5)]
colnames(info)[2]="Condition"
rownames(info)=colnames(count) 

library(DESeq2, quietly = T, warn.conflicts = F)
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)

dds0 <- estimateSizeFactors(dds)

log.norm.counts <- log2(counts(dds0, normalized=TRUE) + 1)
norm.counts <- counts(dds0, normalized=TRUE)
#setwd(outdir)
#write.table(norm.counts, "Adult_DESeq_normalized_data.txt", sep="\t", quote=F, dec=",")

normPCA=t(norm.counts)
normPCA=data.frame(metaAdult[,c(4,9)], normPCA, row.names = metaAdult$Description)
library(FactoMineR, quietly = T, warn.conflicts = F)
res.pca=PCA(normPCA, quali.sup = 1:2, graph=F)
plot.PCA(res.pca, choix="ind", habillage=1, axes=c(1,2), label = "quali")
plot.PCA(res.pca, choix="ind", habillage=2, axes=c(1,2), label = "quali")

dd=dimdesc(res.pca)

#Correlation of illustrative variables to the PCA axe 1
dd$Dim.1$quali
#Correlation of illustrative variables to the PCA axe 2
dd$Dim.2$quali
```

--> The 1st and 2nd axis of the PCA represent about 45% of the total variation and are correlated with the condition.

--> The first axis separate the ablated samples from the samples before ablation and complete regeneration. 

## 2dpa vs 0dpa

```{r, echo=T, message=F, warning=F}
dds0 <- DESeq(dds0)
#resultsNames(dds0)
res20 <- results(dds0, name="Condition_2dpa_vs_0dpa")
#Number of DE genes :
table(res20$padj < 0.05)

setwd(outdir)
nc1=grep("2dpa", metaAdult$Description)
nc2=grep("_0dpa", metaAdult$Description)
res20$Mean1=rowMeans(norm.counts[,nc1])
res20$Mean2=rowMeans(norm.counts[,nc2])
res20$FC=ifelse(res20$Mean1>res20$Mean2,res20$Mean1/res20$Mean2, -res20$Mean2/res20$Mean1)
colnames(res20)[7:8]=c("Mean_2dpa","Mean_0dpa")
#write.table(res20, "Adult_2dpa_vs_0dpa.txt", sep="\t", quote=F, dec=",")
```

## 3dpa vs 2dpa

```{r, echo=T, message=F, warning=F}
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)
dds$Condition <- relevel(dds$Condition, ref = "2dpa")

dds2 <- estimateSizeFactors(dds)
dds2 <- DESeq(dds2)
#resultsNames(dds2)
res23 <- results(dds2, name="Condition_3dpa_vs_2dpa")
print("Number of DE genes :")
table(res23$padj < 0.05)

nc1=grep("3dpa", metaAdult$Description)
nc2=grep("2dpa", metaAdult$Description)
res23$Mean1=rowMeans(norm.counts[,nc1])
res23$Mean2=rowMeans(norm.counts[,nc2])
res23$FC=ifelse(res23$Mean1>res23$Mean2,res23$Mean1/res23$Mean2, -res23$Mean2/res23$Mean1)
colnames(res23)[7:8]=c("Mean_3dpa","Mean_2dpa")
#write.table(res23, "Adult_3dpa_vs_2dpa.txt", sep="\t", quote=F, dec=",")

```


## 10dpa vs 3dpa

```{r, echo=T, message=F, warning=F}
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)
dds$Condition <- relevel(dds$Condition, ref = "3dpa")

dds3 <- estimateSizeFactors(dds)
dds3 <- DESeq(dds3)
#resultsNames(dds3)
res103 <- results(dds3, name="Condition_10dpa_vs_3dpa")
print("Number of DE genes :")
table(res103$padj < 0.05)

nc1=grep("10dpa", metaAdult$Description)
nc2=grep("3dpa", metaAdult$Description)
res103$Mean1=rowMeans(norm.counts[,nc1])
res103$Mean2=rowMeans(norm.counts[,nc2])
res103$FC=ifelse(res103$Mean1>res103$Mean2,res103$Mean1/res103$Mean2, -res103$Mean2/res103$Mean1)
colnames(res103)[7:8]=c("Mean_10dpa","Mean_3dpa")
#write.table(res103, "Adult_10dpa_vs_3dpa.txt", sep="\t", quote=F, dec=",")
```

## 10dpa vs 0dpa

```{r, echo=T, message=F, warning=F}
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)
dds$Condition <- relevel(dds$Condition, ref = "0dpa")

dds10 <- estimateSizeFactors(dds)
dds10 <- DESeq(dds10)
#resultsNames(dds10)
res010 <- results(dds10, name="Condition_10dpa_vs_0dpa")
print("Number of DE genes :")
table(res010$padj < 0.05)

nc1=grep("10dpa", metaAdult$Description)
nc2=grep("0dpa", metaAdult$Description)
res010$Mean1=rowMeans(norm.counts[,nc1])
res010$Mean2=rowMeans(norm.counts[,nc2])
res010$FC=ifelse(res010$Mean1>res010$Mean2,res010$Mean1/res010$Mean2, -res010$Mean2/res010$Mean1)
colnames(res010)[7:8]=c("Mean_10dpa","Mean_0dpa")
write.table(res010, "Adult_10dpa_vs_0dpa.txt", sep="\t", quote=F, dec=",")
```


# DESeq on embryo samples

```{r, message=F, warning=F}
metaEmbryo=metadata[metadata$Stade=="Larvae",]
count=rawcounts[,match(metaEmbryo$Bam_File_Name, colnames(rawcounts))]

info=metaEmbryo[,c(4,5)]
colnames(info)[2]="Condition"
rownames(info)=colnames(count) 

library(DESeq2, quietly = T, warn.conflicts = F)
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)

dds0 <- estimateSizeFactors(dds)

log.norm.counts <- log2(counts(dds0, normalized=TRUE) + 1)
norm.counts <- counts(dds0, normalized=TRUE)
#setwd(outdir)
#write.table(norm.counts, "Embryo_DESeq_normalized_data.txt", sep="\t", quote=F, dec=",")


### PCA on ablated samples (development samples as illustrative samples)
normPCA=t(norm.counts)
normPCA=data.frame(metaEmbryo[,c(4,9)], normPCA, row.names = metaEmbryo$Description)
w=which(metaEmbryo$Condition%in%c("Larvae_80hpf_0hpa", "Larvae_144hpf_0hpa"))

#library(FactoMineR, quietly = T, warn.conflicts = F)
res.pca=PCA(normPCA, quali.sup = 1:2,ind.sup=w,graph=F)
#plot.PCA(res.pca, choix="ind", habillage=2, axes=c(1,2), label = "quali")

res.pca$quali.sup$coord[4,]=colMeans(res.pca$ind.sup$coord[c(2,4,5),])
res.pca$quali.sup$coord[7,]=colMeans(res.pca$ind.sup$coord[c(1,3,6),])

plot.PCA(res.pca, choix="ind", habillage=2, axes=c(1,2), label = "quali")


dd=dimdesc(res.pca, proba=0.1)

#Correlation of illustrative variables to the PCA axe 1
dd$Dim.1$category
#Correlation of illustrative variables to the PCA axe 2
dd$Dim.2$category
```

--> The 1st and 2nd axis of the PCA represent about 49% of the total variation and are correlated with the condition.

--> The first axis separate the ablated samples from the samples before ablation 

--> The second axis separate after regeneration samples from the others. 

## 8hpa vs 0hpa

```{r, echo=T, message=F, warning=F}
dds0 <- DESeq(dds0)
#resultsNames(dds0)
res80 <- results(dds0, name="Condition_8hpa_vs_0hpa")
#Number of DE genes :
table(res80$padj < 0.05)

setwd(outdir)
nc1=grep("8hpa", metaEmbryo$Description)
nc2=grep("0hpa", metaEmbryo$Description)
res80$Mean1=rowMeans(norm.counts[,nc1])
res80$Mean2=rowMeans(norm.counts[,nc2])
res80$FC=ifelse(res80$Mean1>res80$Mean2,res80$Mean1/res80$Mean2, -res80$Mean2/res80$Mean1)
colnames(res80)[7:8]=c("Mean_8hpa","Mean_0hpa")
#write.table(res80, "Embryo_8hpa_vs_0hpa.txt", sep="\t", quote=F, dec=",")
```


## 8hpa vs 80hpf

```{r, echo=T, message=F, warning=F}
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)
dds$Condition <- relevel(dds$Condition, ref = "80hpf")

dds2 <- estimateSizeFactors(dds)
dds2 <- DESeq(dds2)

#resultsNames(dds0)
res880 <- results(dds2, name="Condition_8hpa_vs_80hpf")
#Number of DE genes :
table(res880$padj < 0.05)

setwd(outdir)
nc1=grep("8hpa", metaEmbryo$Description)
nc2=grep("80hpf", metaEmbryo$Description)
res880$Mean1=rowMeans(norm.counts[,nc1])
res880$Mean2=rowMeans(norm.counts[,nc2])
res880$FC=ifelse(res880$Mean1>res880$Mean2,res880$Mean1/res880$Mean2, -res880$Mean2/res880$Mean1)
colnames(res880)[7:8]=c("Mean_8hpa","Mean_80hpf")
#write.table(res880, "Embryo_8hpa_vs_80hpf.txt", sep="\t", quote=F, dec=",")
```



## 3jpa vs 8hpa

```{r, echo=T, message=F, warning=F}
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)
dds$Condition <- relevel(dds$Condition, ref = "8hpa")

dds2 <- estimateSizeFactors(dds)
dds2 <- DESeq(dds2)
#resultsNames(dds2)
res38 <- results(dds2, name="Condition_3jpa_vs_8hpa")
print("Number of DE genes :")
table(res38$padj < 0.05)

nc1=grep("3jpa", metaEmbryo$Description)
nc2=grep("8hpa", metaEmbryo$Description)
res38$Mean1=rowMeans(norm.counts[,nc1])
res38$Mean2=rowMeans(norm.counts[,nc2])
res38$FC=ifelse(res38$Mean1>res38$Mean2,res38$Mean1/res38$Mean2, -res38$Mean2/res38$Mean1)
colnames(res38)[7:8]=c("Mean_3jpa","Mean_8hpa")
#write.table(res38, "Embryo_3jpa_vs_8hpa.txt", sep="\t", quote=F, dec=",")

```


## 3jpa vs 6jpf

```{r, echo=T, message=F, warning=F}
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)
dds$Condition <- relevel(dds$Condition, ref = "6jpf")

dds3 <- estimateSizeFactors(dds)
dds3 <- DESeq(dds3)
#resultsNames(dds3)
res63 <- results(dds3, name="Condition_3jpa_vs_6jpf")
print("Number of DE genes :")
table(res63$padj < 0.05)

nc1=grep("3jpa", metaEmbryo$Description)
nc2=grep("6jpf", metaEmbryo$Description)
res63$Mean1=rowMeans(norm.counts[,nc1])
res63$Mean2=rowMeans(norm.counts[,nc2])
res63$FC=ifelse(res63$Mean1>res63$Mean2,res63$Mean1/res63$Mean2, -res63$Mean2/res63$Mean1)
colnames(res63)[7:8]=c("Mean_3jpa","Mean_6jpf")
#write.table(res63, "Embryo_3jpa_vs_6jpf.txt", sep="\t", quote=F, dec=",")
```


## 3jpa vs 0hpa

```{r, echo=T, message=F, warning=F}
dds=DESeqDataSetFromMatrix(countData=count, colData=info, design=~Lignes+Condition)
dds$Condition <- relevel(dds$Condition, ref = "0hpa")

dds3 <- estimateSizeFactors(dds)
dds3 <- DESeq(dds3)
#resultsNames(dds3)
res30 <- results(dds3, name="Condition_3jpa_vs_0hpa")
print("Number of DE genes :")
table(res30$padj < 0.05)

nc1=grep("3jpa", metaEmbryo$Description)
nc2=grep("0hpa", metaEmbryo$Description)
res30$Mean1=rowMeans(norm.counts[,nc1])
res30$Mean2=rowMeans(norm.counts[,nc2])
res30$FC=ifelse(res30$Mean1>res30$Mean2,res30$Mean1/res30$Mean2, -res30$Mean2/res30$Mean1)
colnames(res30)[7:8]=c("Mean_3jpa","Mean_0hpa")
#write.table(res30, "Embryo_3jpa_vs_0hpa.txt", sep="\t", quote=F, dec=",")
```
