---
title: "WGCNA"
author: "Pauline J"
date: "23 juillet 2018"
output: html_document
---

##  Step 1: Data input and cleaning

### Load library

```{r import library, message=F, warning=F}

setwd("J:\\Pauline J\\Pauline\\Fin_regeneration_project")

##source("http://bioconductor.org/biocLite.R") 
##biocLite(c("AnnotationDbi", "impute", "GO.db", "preprocessCore")) 
##install.packages("WGCNA")
library(impute)
library(WGCNA)
```

### Import file

```{r}
options(stringsAsFactors = FALSE)

AdultData=read.table("Adult_DESeq_normalized_data.txt", sep="\t", dec=",", header=T)
EmbryoData=read.table("Embryo_DESeq_normalized_data.txt", sep="\t", dec=",", header=T)
Metadata=read.table("metadata.txt", header=T, sep="\t")
Metadata$Bam_File_Name=gsub("-",".",Metadata$Bam_File_Name)

wA=which(rowSums(AdultData)==0)
wE=which(rowSums(EmbryoData)==0)
w=wA[wA%in%wE]

AdultData=AdultData[-w,]
EmbryoData=EmbryoData[-w,]

colnames(AdultData)=Metadata$Description[match(colnames(AdultData),Metadata$Bam_File_Name)]
colnames(EmbryoData)=Metadata$Description[match(colnames(EmbryoData),Metadata$Bam_File_Name)]
```

### Load expression data into a multi-set format

```{r}
# Number of sets:
nSets = 2;
# Descriptive names of the two sets.
setLabels = c("Adult Zebrafish", "Embryo Zebrafish")
shortLabels = c("Adult", "Embryo")
# Form multi-set expression data
multiExpr = vector(mode = "list", length = nSets)
multiExpr[[1]] = list(data = as.data.frame(t(AdultData)));
names(multiExpr[[1]]$data) = rownames(AdultData);
rownames(multiExpr[[1]]$data) = names(AdultData);
multiExpr[[2]] = list(data = as.data.frame(t(EmbryoData)));
names(multiExpr[[2]]$data) = rownames(EmbryoData);
rownames(multiExpr[[2]]$data) = names(EmbryoData);
# Check that the data has the correct format for many functions operating on multiple sets:
exprSize = checkSets(multiExpr)

```

### Check that all genes and samples have sufficiently low numbers of missing values.

```{r}
gsg = goodSamplesGenesMS(multiExpr, verbose = 3);

#Keep only good genes
multiExpr[[1]]$data=multiExpr[[1]]$data[gsg$goodGenes]
multiExpr[[2]]$data=multiExpr[[2]]$data[gsg$goodGenes]
```


### Cluster the samples on their Euclidean distance, separately in each set.

```{r}
sampleTrees = list()
for (set in 1:nSets){
sampleTrees[[set]] = hclust(dist(multiExpr[[set]]$data), method = "average")
}

for (set in 1:nSets){
  plot(sampleTrees[[set]], main = paste("Sample clustering on all genes in", setLabels[set]),
       xlab="", sub="", cex = 0.7)
}

```

### Form a multi-set structure that will hold the clinical traits.

```{r}
Traits = vector(mode="list", length = nSets);
for (set in 1:nSets){
  setSamples = rownames(multiExpr[[set]]$data);
  traitRows = match(setSamples, Metadata$Description);
  Traits[[set]] = list(data = Metadata[traitRows, 5:8]);
  rownames(Traits[[set]]$data) = Metadata[traitRows, 1];
}
collectGarbage();
# Define data set dimensions
nGenes = exprSize$nGenes;
nSamples = exprSize$nSamples;


save(multiExpr, Traits, nGenes, nSamples, setLabels, shortLabels, exprSize,
     file = "Consensus-dataInput.RData")

```


## Network construction and module detection

### Choosing the soft-thresholding power: analysis of network topology

```{r}
# Choose a set of soft-thresholding powers
powers = c(seq(4,10,by=1), seq(12,20, by=2));
# Initialize a list to hold the results of scale-free analysis
powerTables = vector(mode = "list", length = nSets);
# Call the network topology analysis function for each set in turn
for (set in 1:nSets){
  powerTables[[set]] = list(data = pickSoftThreshold(multiExpr[[set]]$data, powerVector=powers,
                                                     verbose = 2)[[2]]);
}
collectGarbage();
# Plot the results:
colors = c("black", "red")
# Will plot these columns of the returned scale free analysis tables
plotCols = c(2,5,6,7)
colNames = c("Scale Free Topology Model Fit", "Mean connectivity", "Median connectivity",
             "Max connectivity");

# Get the minima and maxima of the plotted points
ylim = matrix(NA, nrow = 2, ncol = 4);
for (set in 1:nSets){
  for (col in 1:length(plotCols)){
    ylim[1, col] = min(ylim[1, col], powerTables[[set]]$data[, plotCols[col]], na.rm = TRUE);
    ylim[2, col] = max(ylim[2, col], powerTables[[set]]$data[, plotCols[col]], na.rm = TRUE);
  }
}
# Plot the quantities in the chosen columns vs. the soft thresholding power
sizeGrWindow(8, 6)

par(mfcol = c(2,2));
par(mar = c(4.2, 4.2 , 2.2, 0.5))
cex1 = 0.7;
for (col in 1:length(plotCols)) for (set in 1:nSets){
  if (set==1){
    plot(powerTables[[set]]$data[,1], -sign(powerTables[[set]]$data[,3])*powerTables[[set]]$data[,2],
         xlab="Soft Threshold (power)",ylab=colNames[col],type="n", ylim = ylim[, col],
         main = colNames[col]);
    addGrid();
  }
  if (col==1){
    text(powerTables[[set]]$data[,1], -sign(powerTables[[set]]$data[,3])*powerTables[[set]]$data[,2],
    labels=powers,cex=cex1,col=colors[set]);
    legend("bottomright", legend = setLabels, col = colors, pch = 20) ;
  }else{
    text(powerTables[[set]]$data[,1], powerTables[[set]]$data[,plotCols[col]],
    labels=powers,cex=cex1,col=colors[set]);
    legend("topright", legend = setLabels, col = colors, pch = 20) ;
  }
}


```

###  Network construction and consensus module detection

```{r}
net = blockwiseConsensusModules(multiExpr, power = 10, minModuleSize = 30, deepSplit =2,
                                pamRespectsDendro = FALSE,
                                mergeCutHeight = 0.25, numericLabels = TRUE,
                                minKMEtoStay = 0,saveTOMs = TRUE, verbose = 5)

consMEs = net$multiMEs;
moduleLabels = net$colors;
# Convert the numeric labels to color labels
moduleColors = labels2colors(moduleLabels)
consTree = net$dendrograms[[1]];

sizeGrWindow(8,6);

plotDendroAndColors(consTree, moduleColors[net$blockGenes[[1]]],"Module colors",dendroLabels = FALSE, 
                    hang = 0.03,addGuide = TRUE, guideHang = 0.05,
                    main = "Consensus gene dendrogram and module colors")
```

